Demes: a specification for interchanging complex demography

### 2022-03-07 paper.tex
Add text re: the assumed model being Wright-Fisher (#33)

Population sizes are given as the number of individuals.
We focus on the number of individuals as opposed to the number of genomes
in order to not limit other software (need term -- "consumers"?) by placing potentially arbitrary
restrictions on how the numbers of chromosomes, etc., vary amongst individuals.
\jkcomment{As opposed to genomes. Why did we do this?}
\aprcomment{With this, we might want to be able to specify ploidy in as
a graph-level option, which defaults to 2. That would make the reasons
behind this decision obvious then, I think.}
\krtcomment{The number of genomes can be a non-obvious function of the number of individuals, too.
Polymorphism for chromosome number, etc., are all real complications.}
\krtcomment{We support limited aspects of the mating system--more on this below.}
Sizes and mating system details are specified for each population within
epochs. Epochs are contiguous time intervals that define

Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

@@

This model is, by design, very limited. We do not specify any details of
individuals or of the processes that happen \emph{within} populations,
because such details are necessarily complicated, vary greatly by
application, and any attempt to encompass such a broad range of biological
processes is doomed to failure. An ``individual'' (i.e., a discrete
organism) is an abstraction that should have a well defined interpretation
in most cases, despite the endless creativity of biological reality.
This narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. We could never hope to fully encompass all the details
of every possible simulation (including mating system, ploidy,
genome sizes, recombination, mutation and gene conversion rates,
selection coefficients, and so on) and so we clearly demarcate the
role of the specification by including \emph{none} of them.
Such details can be specified by existing mechanisms used by
simulators. If they vary according to the demography (for example,
we have different recombination maps across populations), then
we can unambiguously define these by referring to the population
identifiers and times defined by the Demes model. Rather than attempting
to \emph{embed} this information into the demographic model, we
can define it externally and refer \emph{to} the demography.

@@

>>>>

Population sizes are given as the number of individuals, and details
such as ploidy levels are considered external to the model.
We focus on the number of individuals as opposed to the number of genomes
Sizes and mating system details are specified for each population within
epochs.
Epochs are contiguous time intervals that define
the existence interval of the population. Each epoch specifies the population size
over that interval, which can be a constant value or function defined by start
and end sizes that must remain positive.

\krtcomment{The paragraph(s) below subsume(s)/replace(s) some of what appears in later pargraphs.
	I'm not making an attempt yet to revise that later content--I think other authors should
	weigh in.}

\krtcomment{Aaron -- I have perhaps over-simplified the description of marginal allele frequency dynamics?}

Within a population, we implicitly assume that allele frequency dynamics can
be described by the Wright-Fisher model.  Briefly, generations are non-overlapping (all parents
reproduce and die simultaneously) and the expected frequency (at birth) of the $i^{th}$ allele currently at frequency $p_i$
one generation in the next generation is $p_iw_i/\bar{w}$, where $w_i$ and $\bar{w}$ are the marginal and
mean fitnesses, respectively, properly weighted according to ancestry proportions. In this framework,
a forward-time simulation of finite populations is equivalent to multinomial sampling of allele frequencies each
generation (\citet[][pp 29-31]{burger2000-ul}, \citet[][pp 179-181]{crowkimura1970})
and a backwards-time (coalescent) simulation follows the approximations described in \cite[][chapter 3]{wakeley2008-hd}.
Further, this model assumes "soft" selection \citep{christiansen1975hard}, meaning that the dynamics of population sizes
changes are \textit{independent} of the details of individual fitnesses.
In the \texttt{Discussion}, we discuss alternative models compatible with this framework and outline the expectations for tools implementing
them.
A plethora of forwards and backwards time simulators currently implement this model (REFS).
Each has their own unique method of specifying the details of selection and demography.
The goal of \Demes\ is to provide a common method for the latter task.

\krtcomment{We need to revisit the paragraph(s) below.  We do not assume individuals are exchangeable, because we do not
	require that all individuals have the same fitness. See, e.g., p 60 of Wakeley for why fitness == exchangeability
	and why that is required for the standard coalescent.}

@@

This model is, by design, very limited. We do not specify any details of
individuals or of the processes that happen \emph{within} populations,
because such details are necessarily complicated, vary greatly by
application, and any attempt to encompass such a broad range of biological
processes is doomed to failure. An ``individual'' (i.e., a discrete
organism) is an abstraction that should have a well defined interpretation
in most cases.
This narrow focus on individuals and their groupings into populations
should ensure that the Demes standard is stable, and not subject to
continual development as more and more elaborations of basic processes
are added. We could never hope to fully encompass all the details
of every possible simulation (including ploidy,
genome sizes, recombination, mutation and gene conversion rates,
selection coefficients, and so on) and so we clearly demarcate the
% FIXME Not quite true here, we're including selfing and cloning rate
role of the specification by including \emph{none} of them.
Such details can be specified by existing mechanisms used by
simulators. If they vary according to the demography (for example,
we have different recombination maps across populations), then
we can unambiguously define these by referring to the population
identifiers and times defined by the Demes model. Rather than attempting
to embed this information \emph{in} the demographic model, we
can define it externally and refer \emph{to} the demography.

@@

\krtcomment{Here, we should say something about other models that people may wish to implement within the demes
framework.}
